---
format: noaa-fisheries-memo-to-record-docx
execute: 
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk: 
    dev: ragg_png
    dpi: 300
---

```{r}
library(tidyverse)
library(flextable)
library(scales)
library(janitor)
```

```{r}
noaa_table <-
  function(df) {
    df |>
      flextable::flextable() |>
      flextable::font(fontname = "Times New Roman",
           part = "all") |>
      flextable::fontsize(size = 12) |>
      flextable::theme_zebra() |>
      flextable::autofit()
  }
```


```{r}
survival_by_year <- 
  read_csv("_extensions/noaa-fisheries-memo-to-record/survival_by_year.csv")

vwm <- function(s, V) { # function for variance-weighted mean
  inv.var <- 1 / V # survival rate and the variance of that
  w <- inv.var / sum(inv.var) # survival rate
  s.mean <- sum(s * w)
  s.var <- 1 / sum(inv.var)
  
  list(
    surv = s.mean,
    v.surv = s.var
  )
}

calculate_jpe_data <- function(
    JPI,
    V.JPI,
    timing
) {
  
  f.hat <- 0.4946
  V.f <- 0.006719416
  
  z <- survival_by_year
  
  ## Natural-origin
  s.nat <- z$surv.salt.sac # CJS-estimated surv rates (salt-sac)
  V.s.nat <- z$SE.salt.sac^2 # variances of s.nat
  
  s.nat.mean <- vwm(s.nat, V.s.nat)$surv
  s.nat.var <- vwm(s.nat, V.s.nat)$v.surv
  round.s.nat.mean <- round(s.nat.mean, 4) # force to 4 digits after decimal
  
  ## Hatchery-origin
  s.hat <- z$surv.redding.sac # CJS-estimated surv rates (redding-sac)
  V.s.hat <- z$SE.redding.sac^2 # variances of s.hat
  
  s.hat.mean <- vwm(s.hat, V.s.hat)$surv
  s.hat.var <- vwm(s.hat, V.s.hat)$v.surv
  
  ## ==============================================================================
  ## Forecast natural-origin JPE
  
  JPE <- JPI * f.hat * round(s.nat.mean, 4) # JPE forecast
  # enforce 4 sig digits for surv rate
  
  V.JPE <- (JPI^2 - V.JPI) * # JPE variance (Gray 1999)
    ((f.hat^2 * s.nat.var) + (round.s.nat.mean^2 * V.f) - (V.f * s.nat.var)) +
    (f.hat^2 * round.s.nat.mean^2 * V.JPI)
  
  sd.JPE <- sqrt(V.JPE) # JPE standard deviation
  
  lci <- JPE - (1.96 * sd.JPE) # lower 95% confidence interval
  uci <- JPE + (1.96 * sd.JPE) # upper 95% confidence interval
  
  tibble::tibble(
    JPI = JPI,
    `natural surv rate` = round(s.nat.mean, 4),
    `V natural surv rate` = s.nat.var,
    f.hat = round(f.hat, 4),
    JPE = round(JPE, 0),
    `V JPE` = round(V.JPE),
    `sd JPE` = round(sd.JPE),
    `hatchery surv rate` = round(s.hat.mean, 4),
    `V hatchery surv rate` = s.hat.var,
    `JPE lower 95% CI` = round(lci, 0),
    `JPE upper 95% CI` = round(uci, 0)
  ) |>
    tidyr::pivot_longer(cols = tidyr::everything(),
                        names_to = "Variable") |>
    dplyr::mutate(timing = timing)
}
```


```{r}
options(scipen = 999)
```


**DATE:** [TODO]

**MEMORANDUM FOR:** [TODO]

**FROM:** [TODO]

**SUBJECT:** [TODO]

**ASSOCIATED DOCUMENTS OR DATA FILES:** [TODO]

## SUMMARY

[TODO]

### Variance-weighted Mean Survival

[TODO]

### Confidence Intervals for Juvenile Production Estimate

[TODO]


```{r}
jpe_data_preliminary <-
  calculate_jpe_data(
    JPI = 278448,
    V.JPI = 280340342,
    timing = "Preliminary"
  )

jpe_data_final <-
  calculate_jpe_data(
    JPI = 311058,
    V.JPI = 452638929,
    timing = "Final"
  )

jpe_data_preliminary |>
  bind_rows(jpe_data_final) |>
  pivot_wider(
    id_cols = Variable,
    names_from = timing,
    values_from = value
  ) |>
  filter(Variable %in% c(
    "f.hat",
    "natural surv rate",
    "JPI",
    "JPE",
    "JPE lower 95% CI",
    "JPE upper 95% CI"
  )) |>
  # slice(2, 1, 3, 4, 5) |> # Change if you want to reorder rows
  noaa_table() |> 
  colformat_double(
    i = c(1, 4, 5, 6),
    j = c(2, 3),
    digits = 0,
    big.mark = ","
  )
```


## VWM survival and natural-origin JPE[^1]

[^1]: The Juvenile Production Estimate (JPE) reported here, and implemented in the package WCR.Fisheries.xxxx (available at <NMFS internal github>), uses the "Method 2" approach described in O'Farrell et al. 2018. Staff at the NOAA Fisheries Southwest Fisheries Science Center provide the latest estimates each year for outmigration survival (Arnold Amman) and the fry-to-smolt survival rate (Michael O'Farrell). O’Farrell M. R., W. H. Satterthwaite, A. N. Hendrix, and M. S. Mohr. 2018. Alternative Juvenile Production Estimate (JPE) forecast approaches for Sacramento River winter-run Chinook Salmon. San Francisco Estuary & Watershed Science 16(4):4.  https://doi.org/10.15447/sfews.2018v16iss4art4

### Preliminary estimate

```{r}
jpe_data_preliminary |>
  select(-timing) |>
  rename(`Preliminary Value` = value) |>
  noaa_table() |>
  colformat_double(
    i = c(1, 5, 6, 7, 10, 11),
    j = 2,
    digits = 0,
    big.mark = ","
  ) |>
  colformat_double(
    i = c(2, 4, 8, 9),
    j = 2,
    digits = 4,
    big.mark = ","
  )
```


### Final estimate

```{r}
jpe_data_final |>
  select(-timing) |>
  rename(`Final Value` = value) |>
  noaa_table() |>
  colformat_double(
    i = c(1, 5, 6, 7, 10, 11),
    j = 2,
    digits = 0,
    big.mark = ","
  ) |>
  colformat_double(
    i = c(2, 4, 8, 9),
    j = 2,
    digits = 4,
    big.mark = ","
  )
```


## Exploration of survival by year

```{r}
survival_by_year_tidy <-
  survival_by_year |>
  pivot_longer(cols = -year) |>
  separate_wider_delim(
    name,
    delim = ".",
    names = c(
      "measure",
      "location",
      "region"
    )
  ) |>
  filter(measure != "SE") |>
  mutate(location = str_to_title(location)) |>
  mutate(year = as.character(year))
```


```{r}
#| fig-height: 7
survival_by_year_tidy |>
  ggplot(aes(
    x = year,
    y = value,
    group = location
  )) +
  geom_col() +
  scale_y_continuous(
    labels = percent_format()
  ) +
  facet_wrap(
    vars(location),
    ncol = 1
  ) +
  labs(title = "Survival Rate by Location") +
  theme_minimal(
    base_family = "Arial",
    base_size = 12
  ) +
  theme(
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_text(size = 10),
    plot.title = element_text(
      size = 16,
      face = "bold"
    ),
    plot.title.position = "plot",
    panel.spacing = unit(50, "pt"),
    strip.text = element_text(
      hjust = 0,
      size = 12
    )
  )
```
